### 2021-10-25 text/outline/outline.tex
finish draft of thresholding theory section (+ appendix)

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{/Users/timbarry/optionFiles/mymacros}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

@@

\subsection{Theoretical analysis}

@@

\begin{equation}\label{theoretical_model}
\begin{cases}
m_i = \beta^m_0 + \beta^m_1 p_i + \ep_i \\
g_i = \beta^g_0 + \beta^g_1 p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
For a given threshold $c \in \R$, the imputed perturbation assignment $\hat{p}_i$ is $\hat{p}_i = \mathbb{I}(g_i \geq c).$ The thresholding estimator $\hat{\beta}^m_1$ for $\beta^m_1$ is given by $$\hat{\beta}^m_1 = \frac{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}}) (m_i - \overline{m})}{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})^2 }.$$

\begin{proposition}
 The almost sure limit (as $n \to \infty$) of $\hat{\beta}^m_1$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{a.s.} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi , \\
\omega = \Phi\left(\beta_1^g + \beta_0^g -c \right) ,\\ \zeta = \Phi\left( \beta^g_0 - c \right). \\
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma(\beta^g_1, \pi, c, \beta^g_0) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{a.s.} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^5 \to \R$ be the asymptotic bias of $\hat{\beta}^m_1$, i.e.
$$
b(\beta^g_1, \pi, c, \beta^g_0, \beta^m_1) = \lim_{n\to\infty} \left(\beta^m_1 - \E[\hat{\beta}^m_1]\right) = \beta^m_1 - \E \left( \lim_{a.s.} \hat{\beta}^m_1 \right) = \beta^m_1 - \lim_{a.s.} \hat{\beta}^m_1,
$$
where the $\lim_{a.s.}$ denotes a.s.\ convergence. \textcolor{red}{(HOW TO MAKE THE ABOVE RIGOROUS?)} We can express $b$ in terms of $\gamma$ as $$b(\beta^g_1, \pi, c, \beta^g_0, \beta^m_1) = \beta^m_1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)  \beta^m_1 = \beta^m_1 \left[1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)  \right].$$ The asymptotic bias vanishes when the attenuation function equals $1$.

@@

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we set the perturbation probability $\pi$ to $1/2$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we state several results on the limiting bias $b$ of $\hat{\beta}^m_1$. We refer throughout to Figure \ref{thresholding_theoretical}a, which displays plots of asymptotic bias versus threshold for different values of $\beta^g_1$; the target of inference, $\beta^m_1$, is fixed to $1$ in these plots.

\begin{itemize}
\item 
\begin{proposition}\label{prop:att_bias} Assume without loss of generality that $\beta^m_1 > 0.$ Fix $\pi = 1/2$. For all $(\beta^g_1, c, \beta^g_0) \in \R^3$,
	$$b(\beta^g_1, 1/2, c, \beta^g_0, \beta^m_1) > 0.$$
\end{proposition}
The thresholding method incurs strict attenuation bias (i.e., the true effect size is \textit{under}estimated) for all choices of the threshold and over all possible model parameters (Figure \ref{thresholding_theoretical}a). Attenuation bias is a common attribute of estimators that ignore measurement in errors-in-variables models \cite{Stefanski2000a}.

\item \begin{proposition}
For $\pi = 1/2$ and given $(\beta^g_1, \beta^g_0, \beta^m_1) \in \R^3$, the Bayes-optimal decision boundary $c_\textrm{bayes}$ is a critical value of the limiting bias function $b$, i.e.
$$ \frac{\partial b}{\partial c}\left(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1\right) = 0.$$
\end{proposition}
The Bayes-optimal decision boundary is an optimum (or possibly a saddle point) of the limiting bias function (Figure \ref{thresholding_theoretical}a, vertical blue lines). Interestingly, $c_\textrm{bayes}$ is in some cases a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left) and in other cases a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}\label{prop:limit}
Fix $\pi = 1/2$. As the threshold $c$ tends to infinity, the asymptotic bias $b$ tends to $\beta_1^m/2$, i.e.
$$\lim_{c \to \infty} b(\beta^g_1, 1/2, c, \beta^g_0, \beta^m_1) = \beta_1^m/2.$$
\end{proposition}
In other words, we always can set the threshold to a large number and attain a bias of $\beta_1^m/2$ (Figure \ref{thresholding_theoretical}a, all panels). This result establishes an upper bound on the bias of thresholded regression (under optimal threshold selection strategy).

 \item The following proposition compares the two threshold selection strategies introduced above (i.e., large number versus Bayes-optimal decision boundary) head-to-head.
 \begin{proposition} Let $\beta^m_1 > 0$. For $\beta^g_1 \in [0, 2\Phi^{-1}(3/4))$, we have that $$b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1) > b(\beta^g_1, 1/2, \infty, \beta^g_0, \beta^m_1).$$ For $\beta^g_1 = 2\Phi^{-1}(3/4)$, we have that $$ b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1) = b(\beta^g_1, 1/2, \infty, \beta^g_0, \beta^m_1).$$ Finally, for $\beta^g_1 \in (2\Phi^{-1}(3/4), \infty)$, we have that
 $$b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0, \beta^m_1) < b(\beta^g_1, 1/2, \infty, \beta^g_0, \beta^m_1).$$
 \end{proposition}
That is, setting the threshold to a large number yields a smaller bias when $\beta^g_1$ is small (i.e., $\beta^g_1 < 2\Phi^{-1}(3/4) \approx 1.35$; Figure \ref{thresholding_theoretical}a, left); setting the threshold to the Bayes-optimal decision boundary yields a smaller bias when $\beta^g_1$ is big (i.e., $\beta^g_1 > 2\Phi^{-1}(3/4)$; Figure \ref{thresholding_theoretical}a, right); and the two approaches coincide when $\beta^g_1$ is intermediate (i.e., $\beta^g_1 = 2\Phi^{-1}(3/4)$; Figure \ref{thresholding_theoretical}a, middle).

\item
 \begin{proposition}\label{prop_monotonic} Fix $\pi = 1/2$. The limiting bias $b$ decreases monotonically in $\beta_1^g$, i.e.
$$\frac{\partial b}{\partial(\beta^g_1)}\left(\beta^g_1, 1/2, c, \beta^g_0, \beta^m_1\right) \leq 0.$$
\end{proposition}
This result formalizes the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated. To visualize Proposition (\ref{prop_monotonic}), one can fix a threshold (e.g., $c = 0$) and scan for bias across the panels.
\end{itemize}

@@

\begin{center}
\textbf{Generalizing to $\pi \in [0,1/2]$ (Panel b)}
\end{center}
We generalize the expression for bias to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}
As the threshold $c$ tends to infinity, the limiting bias $b$ tends to $\beta^m_1 \pi,$ i.e.
$$ \lim_{ c \to \infty } b(\beta^g_1, \pi, c, \beta^g_0, \beta^m_1) = \beta^m_1 \pi.$$
\end{proposition}
Thus, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the asymptotic bias is $\beta^m_1 \pi$ (Figure \ref{thresholding_theoretical}b; $\beta^m_1$ fixed at $1$ in plot). We can understand this result intuitively by considering an extreme example: when $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/thresholding_theoretical/plot}
	\caption{\textbf{Theoretical challenges of thresholded regression.} \textbf{a,} (Asymptotic) bias versus threshold for different values of $\beta^g_1$. The bias function is highly nonconvex and strictly nonzero. Across all panels, $\beta^g_0 = 0, \beta^m_1 = 1$, and $\pi = 1/2$. \textbf{b,} Bias versus $\pi$ when the threshold is set to a large number. Bias is directly proportional to $\pi$ in this setting. $\beta^m_1 = 1$. \textbf{c,} Bias-variance decomposition in no-intercept model. Bias decreases and variance increases as the threshold tends to infinity. $\beta^m_1 = 1, \beta^g_1 = 1,$ and $\pi = 0.1$.}
	\label{thresholding_theoretical}
\end{figure}

@@

Finally, to shed light on limitations of the large threshold selection strategy, we derive an exact bias-variance decomposition for the thresholding estimator. We consider a slightly simpler, no-intercept version of (\ref{theoretical_model}) for this purpose:
\begin{equation}\label{theoretical_model_no_int}
\begin{cases}
m_i = \beta_m p_i + \ep_i \\
g_i = \beta_g p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
The thresholding estimator $\hat{\beta}_m$ in the no-intercept case is
\begin{equation}\label{thresh_estimator_no_int}
 \hat{\beta}_m = \frac{ \sum_{i=1}^n \hat{p}_i m_i }{\sum_{i=1}^n \hat{p}_i^2}. 
\end{equation}
\begin{proposition}
The limiting distribution of $\hat{\beta}_m$ is
$$\sqrt{n}(\hat{\beta}_m - l) \xrightarrow{d} N\left(0, \frac{ \beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2) }{\left(\E[\hat{p}_i]\right)^2} \right),$$ where
$$\begin{cases}
l = \beta_m\omega \pi/[\zeta(1-\pi) + \omega \pi], \\
\E[\hat{p}_i] = \pi \omega + (1-\pi) \zeta, \\
\omega = \Phi(\beta_g - c), \\
\zeta = \Phi(- c).\\
\end{cases}
$$
\end{proposition}
This result yields an exact bias-variance decomposition for $\hat{\beta}_m$ for large $n$ (Figure \ref{thresholding_theoretical}c). As the threshold tends to infinity, the bias decreases and the variance increases, consistent with the intuition that a large threshold reduces the misclassification rate at the cost of decreasing the ``effective sample size.'' The best strategy for maximizing estimation accuracy (as quantified by mean squared error) is to select a threshold that induces moderate bias. A downside of this approach, however, is that constructing valid confidence intervals becomes more challenging.

@@

\subsection{Theoretical details for thresholding estimator}

This section contains proofs for results presented the section ``Theoretical analysis of thresholding estimator.'' The subsections are arranged as follows: FILL IN...

\subsubsection{Notation}
 All notation introduced here pertains to the Gaussian model with intercepts (\ref{theoretical_model}). Recall that the attenuation function $\gamma: \R^4 \to \R$ is defined by
$$ \gamma(\beta^g_1, c, \pi, \beta^g_0) = \frac{\pi(\omega - \E[\hat{p}_i])}{ \E[\hat{p}_i](1 -\E[\hat{p}_i])},$$ where $$\begin{cases}
\omega = \Phi\left(c - \beta_1^g - \beta_0^g \right) ,\\ \zeta = \Phi\left( c - \beta^g_0 \right), \\ \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi. 
\end{cases}$$ Additionally, recall that the bias function $b: \R^5 \to \R$ is defined by
$$ b(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1) = \beta^m_1 (1 - \gamma(\beta^g_1, c, \pi, \beta^g_0)).$$ The parameters $\beta^g_0$ and $\beta^m_1$ are given, fixed constants throughout the proofs. Therefore, to minimize notation, we typically use $\gamma(\beta^g_1, c, \pi)$ (resp., $b(\beta^g_1, c, \pi)$) to refer to the function $\gamma$ (resp., $b$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$ (resp., $(\beta^g_1, c, \pi, \beta^g_0, \beta^m_1)$). Next, we define the functions $g$ and $h: \R^4 \to \R$ by
\begin{equation}\label{def_g}
g(\beta^g_1, c, \pi, \beta^g_0) = (1-\pi)\left( \int_{ -\infty}^{\beta_0^g + \beta_1^g - c} f \right) - (1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right)\end{equation}
 and
\begin{multline}\label{def_h}
h(\beta^g_1, c, \pi, \beta^g_0) = \left[(1-\pi)\left( \int_{-\infty}^{ \beta_0^g - c} f \right) + \pi\left( \int_{-\infty}^{ \beta^g_0 + \beta^g_1 - c} f \right) \right] \cdot \\ \left[(1-\pi)\left( \int_{-\infty}^{c - \beta^g_0} f \right) + \pi\left(\int_{-\infty}^{c - \beta_0^g - \beta_1^g} f \right) \right],
\end{multline} where $f: \R \to [0,\infty)$ is the $N(0,1)$ density. We will refer to $g$ and $h$ repeatedly throughout the proofs. Similarly to above, we typically use the notation $g(\beta^g_1, c, \pi)$ (resp.\ $h(\beta^g_1, c, \pi)$) to refer to $g$ (resp.\ $h$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$.

Finally, for a given function $r: \R^{p} \to \R$, point $x \in \R^p$, and index $i \in \{1, \dots, p\}$, we use the symbol $D_i r(x)$ to refer to the derivative of the $i$th component of $r$ evaluated at $x$ (\textit{sensu} \cite{fitzpatrick2009}). For example, $D_1 \gamma(\beta^g_1, c, 1/2)$ is the derivative of the first component of $\gamma$ (the component corresponding to $\beta^g_1$) evaluated at $(\beta^g_1, c, 1/2)$. Likewise,  $D_2g(\beta^g_1, c, \pi)$ is the derivative of the second component of $g$ (the component corresponding to $c$) evaluated at $(\beta^g_1, c, \pi).$

\subsubsection{Limit of $\hat{\beta}^m_1$}

>>>>

\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\setcounter{tocdepth}{2}
\usepackage{xcolor}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\usepackage{float}
\usepackage[caption = false]{subfig}
\usepackage{/Users/timbarry/optionFiles/mymacros}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

\tableofcontents

@@

\subsection{Theoretical analysis}\label{sec:thresholding_theory}

@@

\begin{equation}\label{theoretical_model}
\begin{cases}
m_i = \beta^m_0 + \beta^m_1 p_i + \ep_i \\
g_i = \beta^g_0 + \beta^g_1 p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
For a given threshold $c \in \R$, the imputed perturbation assignment $\hat{p}_i$ is given by $\hat{p}_i = \mathbb{I}(g_i \geq c).$ The thresholding estimator $\hat{\beta}^m_1$ for $\beta^m_1$ is $$\hat{\beta}^m_1 = \frac{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}}) (m_i - \overline{m})}{\sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})^2 }.$$

\begin{proposition}\label{prop:convergence}
 The almost sure limit (as $n \to \infty$) of $\hat{\beta}^m_1$ is
\begin{equation}\label{thresh_est_intercepts}
\hat{\beta}^m_1 \xrightarrow{a.s.} \beta^m_1 \left(\frac{ \pi( \omega - \E[ \hat{p}_i ])}{ \E[\hat{p}_i] (1 - \E[\hat{p}_i])}\right),
\end{equation}
where
$$\begin{cases}
\E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi , \\
\omega = \Phi\left(\beta_1^g + \beta_0^g -c \right) ,\\ \zeta = \Phi\left( \beta^g_0 - c \right). \\
\end{cases}$$
\end{proposition}
Let $\gamma: \R^4 \to \R$ be defined by
$$ \gamma(\beta^g_1, \pi, c, \beta^g_0) = \frac{\pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i] (1 - \E[\hat{p}_i]) }.$$ We call $\gamma$ the ``attenuation function.'' Observe that \begin{itemize}
\item[i.] $\gamma$ does not depend on $\beta^m_1$ or $\beta^m_0$, and
\item[ii.] $\hat{\beta}^m_1 \xrightarrow{a.s.} [\gamma(\beta_0^g, \beta_1^g, c, \pi)] \beta^m_1.$
\end{itemize}
Let $b: \R^4 \to \R$ be the asymptotic \textit{relative} bias of $\hat{\beta}^m_1$:
\begin{multline*}
b(\beta^g_1, \pi, c, \beta^g_0) = \left(\frac{1}{\beta^m_1}\right) \lim_{n\to\infty} \left(\beta^m_1 - \E[\hat{\beta}^m_1]\right) = \left( \frac{1}{\beta^m_1}\right)\left(\beta^m_1 - \E \left(\lim_{a.s.} \hat{\beta}^m_1 \right)\right) \\ = \frac{1}{\beta^m_1} \left(\beta^m_1 - \gamma(\beta^g_1, \pi, c, \beta^g_0)\beta^m_1\right) = 1 - \gamma(\beta^g_1, \pi, c, \beta^g_0),
\end{multline*}
where $\lim_{a.s.}$ denotes a.s.\ convergence. The asymptotic relative bias vanishes when the attenuation function equals $1$. \textcolor{red}{(HOW TO MAKE THE ABOVE RIGOROUS?)}

@@

To investigate the basic question of ``What is a good threshold selection strategy?'', we study the relationship between the asymptotic relative bias $b$ of $\hat{\beta}^m_1$ and the selected threshold $c$. For simplicity, we set the perturbation probability $\pi$ to $1/2$. Let $c_\textrm{bayes} \in \R$ be the Bayes-optimal decision boundary for classifying cells as perturbed or unperturbed, i.e. $$c_\textrm{bayes} = \argmin_{c \in \R} \P(\hat{p}_i \neq p_i).$$ Simple algebra shows that $c_\textrm{bayes} = \beta_0^g + (1/2) \beta^g_1.$ Below, we give several results for the asymptotic relative bias $b$ of $\hat{\beta}^m_1$. We refer throughout to Figure \ref{thresholding_theoretical}a, which displays plots of asymptotic relative bias versus threshold for different values of $\beta^g_1$. We sometimes refer to ``asymptotic relative bias'' using the shortened term ``bias'' for succinctness.

\begin{itemize}
\item 
\begin{proposition}\label{prop:att_bias} Fix $\pi = 1/2$. For all $(\beta^g_1, c, \beta^g_0) \in \R^3$, the asymptotic relative bias is positive, i.e. 
$$b(\beta^g_1, 1/2, c, \beta^g_0) > 0.$$
\end{proposition}
The thresholding method incurs strict attenuation bias (i.e., it \textit{under}estimates the true effect size) for all choices of the threshold and over all possible values of the model parameters (Figure \ref{thresholding_theoretical}a). Attenuation bias is a common attribute of estimators that ignore measurement in errors-in-variables models \cite{Stefanski2000a}.
\item
\begin{proposition}\label{prop:monotonic} Fix $\pi = 1/2$. The asymptotic relative bias $b$ decreases monotonically in $\beta_1^g$, i.e.
	$$\frac{\partial b}{\partial(\beta^g_1)}\left(\beta^g_1, 1/2, c, \beta^g_0\right) \leq 0.$$
\end{proposition}
This result formalizes the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated. To visualize Proposition (\ref{prop:monotonic}), one can fix a threshold (e.g., $c = 0$) and scan for bias across the panels.	
\item \begin{proposition}\label{prop:bayes_opt}
For $\pi = 1/2$ and given $(\beta^g_1, \beta^g_0) \in \R^2$, the Bayes-optimal decision boundary $c_\textrm{bayes}$ is a critical value of the bias function $b$, i.e.
$$ \frac{\partial b}{\partial c}\left(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0\right) = 0.$$
\end{proposition}
The Bayes-optimal decision boundary is an optimum (or possibly a saddle point) of the asymptotic relative bias function (Figure \ref{thresholding_theoretical}a, vertical blue lines). Interestingly, $c_\textrm{bayes}$ is in some cases a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left) and in other cases a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}\label{prop:c_limit_half}
Assume without loss of generality that $\beta^g_1 > 0$, and fix $\pi = 1/2$. As the threshold $c$ tends to infinity, the asymptotic relative bias $b$ tends to $1/2$, i.e.
$$\lim_{c \to \infty} b(\beta^g_1, 1/2, c, \beta^g_0) = 1/2.$$
\end{proposition}
In other words, we always can set the threshold to a large number and attain a relative bias of $1/2$ (Figure \ref{thresholding_theoretical}a, all panels). This result establishes an upper bound on the bias of thresholded regression (under optimal threshold selection strategy).

 \item The following proposition compares the two threshold selection strategies introduced above (i.e., large number versus Bayes-optimal decision boundary) head-to-head.
 \begin{proposition}\label{prop:comparison} Assume without loss of generality that $\beta^g_1 > 0$. For $\beta^g_1 \in [0, 2\Phi^{-1}(3/4))$, we have that $$b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0) > b(\beta^g_1, 1/2, \infty, \beta^g_0).$$ For $\beta^g_1 = 2\Phi^{-1}(3/4)$, we have that $$ b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0) = b(\beta^g_1, 1/2, \infty, \beta^g_0).$$ Finally, for $\beta^g_1 \in (2\Phi^{-1}(3/4), \infty)$, we have that
 $$b(\beta^g_1, 1/2, c_\textrm{bayes}, \beta^g_0) < b(\beta^g_1, 1/2, \infty, \beta^g_0).$$
 \end{proposition}
Setting the threshold to a large number yields a smaller bias when $\beta^g_1$ is small (i.e., $\beta^g_1 < 2\Phi^{-1}(3/4) \approx 1.35$; Figure \ref{thresholding_theoretical}a, left); setting the threshold to the Bayes-optimal decision boundary yields a smaller bias when $\beta^g_1$ is large (i.e., $\beta^g_1 > 2\Phi^{-1}(3/4)$; Figure \ref{thresholding_theoretical}a, right); and the two approaches coincide when $\beta^g_1$ is intermediate (i.e., $\beta^g_1 = 2\Phi^{-1}(3/4)$; Figure \ref{thresholding_theoretical}a, middle).
\end{itemize}

@@

\begin{center}
\textbf{Generalizing to $\pi \in [0,1/2]$ (Panel b)}
\end{center}
We generalize the expression for bias when the threshold is large to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}\label{prop:c_limit}
Assume without loss of generality that $\beta^g_1 > 0$. As the threshold $c$ tends to infinity, the asymptotic relative bias $b$ tends to $\pi,$ i.e.
$$ \lim_{ c \to \infty } b(\beta^g_1, \pi, c, \beta^g_0) = \pi.$$
\end{proposition}
In other words, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the asymptotic relative bias is $\pi$ (Figure \ref{thresholding_theoretical}b). We can understand this result intuitively by considering an extreme example: when $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

\begin{figure}[h!]
	\centering
	\includegraphics[width=1\linewidth]{../../figures/thresholding_theoretical/plot}
	\caption{\textbf{Theoretical challenges of thresholded regression.} \textbf{a,} Asymptotic relative bias versus threshold for different values of $\beta^g_1$. The bias function is highly nonconvex and strictly nonzero. Vertical blue lines, Bayes-optimal decision boundaries. Across all panels, $\beta^g_0 = 0$ and $\pi = 1/2$. \textbf{b,} Asymptotic relative bias versus $\pi$ when the threshold is set to a large number. The two quantities coincide exactly. \textbf{c,} Bias-variance decomposition for thresholding method in no-intercept model. Bias decreases and variance increases as the threshold tends to infinity. $\beta^g_1 = 1, \beta^m_1 = 1,$ and $\pi = 0.1$.}
	\label{thresholding_theoretical}
\end{figure}

@@

Finally, to shed light on limitations of the large threshold selection strategy, we derive an exact bias-variance decomposition for the thresholding estimator. We consider a slightly simpler, no-intercept version of (\ref{theoretical_model}) for this purpose:
\begin{equation}\label{theoretical_model_no_int}
\begin{cases}
m_i = \beta_m p_i + \ep_i \\
g_i = \beta_g p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i.
\end{cases}
\end{equation}
The thresholding estimator $\hat{\beta}_m$ in the no-intercept case is
\begin{equation}\label{thresh_estimator_no_int}
 \hat{\beta}_m = \frac{ \sum_{i=1}^n \hat{p}_i m_i }{\sum_{i=1}^n \hat{p}_i^2}. 
\end{equation}
\begin{proposition}\label{prop:bv_decomp}
The limiting distribution of $\hat{\beta}_m$ is
$$\sqrt{n}(\hat{\beta}_m - l) \xrightarrow{d} N\left(0, \frac{ \beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2) }{\left(\E[\hat{p}_i]\right)^2} \right),$$ where
$$\begin{cases}
l = \beta_m\omega \pi/[\zeta(1-\pi) + \omega \pi], \\
\E[\hat{p}_i] = \pi \omega + (1-\pi) \zeta, \\
\omega = \Phi(\beta_g - c), \\
\zeta = \Phi(- c).\\
\end{cases}
$$
\end{proposition}
This result yields an exact bias-variance decomposition for $\hat{\beta}_m$ for large $n$ (Figure \ref{thresholding_theoretical}c). As the threshold tends to infinity, the bias decreases and the variance increases, consistent with the intuition that a large threshold reduces the misclassification rate at the cost of decreasing the ``effective sample size.'' The best strategy for maximizing estimation accuracy (as quantified by mean squared error) is to select a threshold that induces moderate bias. A downside of this approach, however, is that constructing valid confidence intervals becomes more challenging.

@@

\subsection{Theoretical details for thresholding estimator}
This section contains proofs of the propositions presented Section \ref{sec:thresholding_theory}, ``Theoretical analysis of thresholding estimator.'' The subsections are organized as follows. Section (\ref{sec:notation}) introduces some notation. Section (\ref{sec:convergence}) establishes almost sure convergence of the thresholding estimator in the model (\ref{theoretical_model}), proving Proposition \ref{prop:convergence}. Section (\ref{sec:simplication}) simplifies the expression for the attenuation function $\gamma$, and section (\ref{sec:derivatives})  computes derivatives of $\gamma$ to be used throughout the proofs. Section (\ref{sec:c_limit}) establishes the limit in $c$ of $\gamma$, proving Proposition \ref{prop:c_limit} and as a corollary Proposition \ref{prop:c_limit_half}. Section (\ref{sec:bayes_opt}) establishes that the Bayes-optimal decision boundary is a critical value of $\gamma$, proving Proposition \ref{prop:bayes_opt}, and section (\ref{sec:comparison}) compares the competing threshold selection strategies head-to-head, proving Proposition \ref{prop:comparison}. Section (\ref{sec:monotone}) demonstrates that $\gamma$ is monotone in $\beta^g_1$, proving Proposition \ref{prop:monotonic}, and Section (\ref{sec:att_bias}) establishes attenuation bias of the thresholding estimator, proving Proposition \ref{prop:att_bias}. Finally, Section (\ref{sec:bv_decomp}) derives the bias-variance decomposition of the thresholding estimator in the model (\ref{theoretical_model_no_int}), proving Proposition \ref{prop:bv_decomp}.

% Labels of propositions and theorems
% propositions:
% prop:convergence
% prop:att_bias
% prop:bayes_opt
% prop:c_limit_half
% prop:comparison
% prop:monotonic
% prop:c_limit
% prop:bv_decomp

% sections
% sec:notation
% sec:convergence
% sec:simplication
% sec:derivatives
% sec:c_limit
% sec:bayes_opt
% sec:comparison
% sec:beta_lim
% sec:monotone
% sec:att_bias
% sec:bv_decomp

\subsubsection{Notation}\label{sec:notation}
 All notation introduced in this subsection (i.e., \ref{sec:notation}) pertains to the Gaussian model with intercepts (\ref{theoretical_model}). Recall that the attenuation function $\gamma: \R^4 \to \R$ is defined by
$$ \gamma(\beta^g_1, c, \pi, \beta^g_0) = \frac{\pi(\omega - \E[\hat{p}_i])}{ \E[\hat{p}_i](1 -\E[\hat{p}_i])},$$ where $$\begin{cases} \E[\hat{p}_i] = \zeta(1-\pi) + \omega\pi, \\
\omega = \Phi\left(\beta_1^g + \beta_0^g - c \right) ,\\ \zeta = \Phi\left( \beta^g_0 - c \right).
\end{cases}$$ Additionally, recall that the asymptotic relative bias function $b: \R^4 \to \R$ is
$$ b(\beta^g_1, c, \pi, \beta^g_0) = 1 - \gamma(\beta^g_1, c, \pi, \beta^g_0).$$ Next, we define the functions $g$ and $h: \R^4 \to \R$ by
\begin{equation}\label{def_g}
g(\beta^g_1, c, \pi, \beta^g_0) = (1-\pi)\left( \Phi(\beta_0^g + \beta_1^g - c)\right) - (1-\pi)\left(\Phi(\beta_0^g - c)\right)\end{equation}
 and
\begin{multline}\label{def_h}
h(\beta^g_1, c, \pi, \beta^g_0) = \left[(1-\pi)\left( \Phi(\beta_0^g - c)\right) + \pi\left(\Phi(\beta^g_0 + \beta^g_1 - c) \right) \right] \cdot \\ \left[(1-\pi)\left( \Phi(c - \beta^g_0) \right) + \pi\left(\Phi(c - \beta_0^g - \beta_1^g) \right) \right].
\end{multline}
We use $f:\R \to \R$ to denote the $N(0,1)$ density, and we denote the right-tail probability probability of $f$ by $\bar{\Phi}$, i.e., 
$$\bar{\Phi}(x) = \int_{x}^{\infty} f = \Phi(-x).$$

The parameter $\beta^g_0$ is a given, fixed constant throughout the proofs. Therefore, to minimize notation, we typically use $\gamma(\beta^g_1, c, \pi)$ (resp., $b(\beta^g_1, c, \pi),$ $g(\beta^g_1, c, \pi),$ $h(\beta^g_1, c, \pi)$) to refer to the function $\gamma$ (resp., $b, g, h$) evaluated at $(\beta^g_1, c, \pi, \beta^g_0)$. Finally, for a given function $r: \R^{p} \to \R$, point $x \in \R^p$, and index $i \in \{1, \dots, p\}$, we use the symbol $D_i r(x)$ to refer to the derivative of the $i$th component of $r$ evaluated at $x$ (\textit{sensu} \cite{fitzpatrick2009}). For example, $D_1 \gamma(\beta^g_1, c, 1/2)$ is the derivative of the first component of $\gamma$ (the component corresponding to $\beta^g_1$) evaluated at $(\beta^g_1, c, 1/2)$. Likewise,  $D_2g(\beta^g_1, c, \pi)$ is the derivative of the second component of $g$ (the component corresponding to $c$) evaluated at $(\beta^g_1, c, \pi).$

\subsubsection{Almost sure limit of $\hat{\beta}^m_1$}\label{sec:convergence}
